---
title: "Initial Analysis and Model Performance"
author: "Sanatan Das"
date: "April 16, 2018"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center}\LARGE\includegraphics[width=5cm]{logo.png}\\[\bigskipamount]}
- \posttitle{\end{center}}
urlcolor: blue
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"logo.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(broom)
library(knitr)
```

\newpage

# Initial Data Analysis

## Load the Data

```{r var1, eval=TRUE, echo=TRUE}

# load the data set from excel file
default_rates <- read_excel("C:/view/opt/apps/git/compscix-415-1-assignments/data/peps3xx.xls")

```

## The Variables

```{r var2, eval=TRUE, echo=TRUE}

# take a look at the data
glimpse(default_rates)

```

(Data Source : [Federal Student Aid](https://studentaid.ed.gov/sa/about/data-center/student/default))

(Data Definition : [Instructions for Using the Data Files](https://www2.ed.gov/offices/OSFAP/defaultmanagement/instructions.html))

## Problem Category (Regression)

Our target variable is *default rate* or *drate* which is a numerical (double) variable. So our problem is categorically regression problem. We will create different kind of refression models to predict our data.

## Predictors Data Types

```{r var3, eval=TRUE, echo=TRUE}

# add factor to the 'char' columns
default_rates$Name <- as.factor(default_rates$Name)
default_rates$State <- as.factor(default_rates$State)
default_rates$ZipCode <- as.factor(default_rates$ZipCode)
default_rates$ProgLength <- as.factor(default_rates$ProgLength)
default_rates$SchoolType <- as.factor(default_rates$SchoolType)
default_rates$EthnicCode <- as.factor(default_rates$EthnicCode)
default_rates$Prate <- as.factor(default_rates$Prate)
default_rates$CongDis <- as.factor(default_rates$CongDis)
# convert the columns to 'double' data type
default_rates$Drate <- as.double(default_rates$Drate)
default_rates$Num <- as.double(default_rates$Num)
default_rates$Denom <- as.double(default_rates$Denom)

```

# Data Visualization

## Relationship between Ethnic Code and the Default Rate

```{r var4, eval=TRUE, echo=TRUE}

# plot the relationship between Ethnic Code and the Default Rate (Bar Graph)
default_rates %>% 
ggplot() + 
  geom_bar(aes(x = EthnicCode, y = Drate), stat = 'identity') +
  labs(x="Ethnic Code", y="Default Rate") +
  theme_bw()

# plot the relationship between Ethnic Code and the Default Rate (Box Plot)
default_rates %>% 
ggplot(aes(x = EthnicCode, y = Drate), group=EthnicCode) + 
  geom_boxplot() +
  labs(x="Ethnic Code", y="Default Rate") +
  theme_bw()

```

## Relationship between Program Length and the Default Rate

```{r var5, eval=TRUE, echo=TRUE}

# plot the relationship between Program Length and the Default Rate (Bar Graph)
default_rates %>% 
ggplot() + 
  geom_bar(aes(x = ProgLength, y = Drate), stat = 'identity') +
  labs(x="Program Length", y="Default Rate") +
  theme_bw()

# plot the relationship between Prog Length and the Default Rate (Box Plot)
default_rates %>% 
ggplot(aes(x = ProgLength, y = Drate), group=ProgLength) + 
  geom_boxplot() +
  labs(x="Program Length", y="Default Rate") +
  theme_bw()

```

## Relationship between School Type and the Default Rate
```{r var6, eval=TRUE, echo=TRUE}

# plot the relationship between School Type and the Default Rate (Bar Graph)
default_rates %>% 
ggplot() + 
  geom_bar(aes(x = SchoolType, y = Drate), stat = 'identity') +
  labs(x="School Type", y="Default Rate") +
  theme_bw()

# plot the relationship between School Type and the Default Rate (Box Plot)
default_rates %>% 
ggplot(aes(x = SchoolType, y = Drate), group=SchoolType) + 
  geom_boxplot() +
  labs(x="School Type", y="Default Rate") +
  theme_bw()

```

## Relationship between Prate (Type of rate calculated for the institution.) and the Default Rate

```{r var7, eval=TRUE, echo=TRUE}

# plot the relationship between Prate (Type of rate was calculated for the institution) 
# and the Default Rate (Bar Graph)
default_rates %>% 
ggplot() + 
  geom_bar(aes(x = Prate, y = Drate), stat = 'identity') +
  labs(x="Type of rate was calculated for the institution", y="Default Rate") +
  theme_bw()

# plot the relationship between Prate (Type of rate was calculated for the institution) 
# and the Default Rate (Box Plot)
default_rates %>% 
ggplot(aes(x = Prate, y = Drate), group=Prate) + 
  geom_boxplot() +
  labs(x="Type of rate was calculated for the institution", y="Default Rate") +
  theme_bw()

```

## Relationship between Num (Number of Borrowers in Default) and the Default Rate

```{r var8, eval=TRUE, echo=TRUE}

# plot the relationship between Num (Number of Borrowers in Default) 
# and the Default Rate
ggplot(default_rates, aes(x = Num, y = Drate)) +
  geom_point(color = "blue") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(x="Number of Borrowers in Default", y="Default Rate") +
  theme_bw()

```

## Relationship between Denom (Number of Borrowers in Repay) and the Default Rate

```{r var9, eval=TRUE, echo=TRUE}

# plot the relationship between Denom (Number of Borrowers in Repay) 
# and the Default Rate
ggplot(default_rates, aes(x = Denom, y = Drate)) +
  geom_point(color = "blue") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(x="Number of Borrowers in Repay", y="Default Rate") +
  theme_bw()

```

\newpage

# Model Performance (initial naive models)


```{r var10, eval=FALSE, echo=FALSE}

# Linear model on 'Name' feature
lm_1 <- lm(Drate ~ Name, data = default_rates)
tidy(lm_1)

```

```{r var11, eval=FALSE, echo=FALSE}

# Linear model on 'Name' feature
lm_1 <- lm(Drate ~ Name, data = default_rates)
tidy(lm_1)

```

```{r var12, eval=FALSE, echo=FALSE}

# Linear model on 'State' feature
lm_2 <- lm(Drate ~ State, data = default_rates)
tidy(lm_2)

```

```{r var13, eval=FALSE, echo=FALSE}

# Linear model on 'ZipCode' feature
lm_3 <- lm(Drate ~ ZipCode, data = default_rates)
tidy(lm_3)

```

## Model evaluation on 'ProgLength' feature

```{r var14, eval=TRUE, echo=TRUE}

# Linear model on 'ProgLength' feature
lm_4 <- lm(Drate ~ ProgLength, data = default_rates)
glance(lm_4)

```
## Model evaluation on 'SchoolType' feature

```{r var15, eval=TRUE, echo=TRUE}

# Linear model on 'SchoolType' feature
lm_5 <- lm(Drate ~ SchoolType, data = default_rates)
glance(lm_5)

```

## Model evaluation on 'Num' feature

```{r var16, eval=TRUE, echo=TRUE}

# Linear model on 'Num' feature
lm_6 <- lm(Drate ~ Num, data = default_rates)
tidy(lm_6)

```
## Model evaluation on 'Denom' feature

```{r var17, eval=TRUE, echo=TRUE}

# Linear model on 'Denom' feature
lm_7 <- lm(Drate ~ Denom, data = default_rates)
tidy(lm_7)

```

```{r var18, eval=FALSE, echo=FALSE}

# Linear model on 'Prate' feature
lm_8 <- lm(Drate ~ Prate, data = default_rates)
tidy(lm_8)

```

## Model evaluation on 'EthnicCode' feature

```{r var19, eval=TRUE, echo=TRUE}

# Linear model on 'EthnicCode' feature
lm_9 <- lm(Drate ~ EthnicCode, data = default_rates)
glance(lm_9)

```

## Multiple Linear Regression Model

```{r var20, eval=TRUE, echo=TRUE}

# Final model usiing all features
lm_10 <- lm(Drate ~ ProgLength + SchoolType + Num + Denom + Prate + EthnicCode, data = default_rates)
glance(lm_10)

```

