---
title: "Model Performance - Linear Regression"
author: "Sanatan Das"
date: "April 24, 2018"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center}\LARGE\includegraphics[width=5cm]{logo.png}\\[\bigskipamount]}
- \posttitle{\end{center}}
urlcolor: blue
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"logo.png\" style=\"float: right;width: 150px;\"/>')
   });
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(broom)
library(knitr)
```

\newpage

```{r var1, eval=TRUE, echo=FALSE}

# load the data set from excel file
default_rates <- read_excel("C:/view/opt/apps/git/compscix-415-1-assignments/data/peps3xx.xls")

# add factor to the 'char' columns
default_rates$Name <- as.factor(default_rates$Name)
default_rates$State <- as.factor(default_rates$State)
default_rates$ZipCode <- as.factor(default_rates$ZipCode)
default_rates$ProgLength <- as.factor(default_rates$ProgLength)
default_rates$SchoolType <- as.factor(default_rates$SchoolType)
default_rates$EthnicCode <- as.factor(default_rates$EthnicCode)
default_rates$Prate <- as.factor(default_rates$Prate)
default_rates$CongDis <- as.factor(default_rates$CongDis)
# convert the columns to 'double' data type
default_rates$Drate <- as.double(default_rates$Drate)
default_rates$Num <- as.double(default_rates$Num)
default_rates$Denom <- as.double(default_rates$Denom)

```

# Basic Models (Linear Regression)

## Model evaluation on 'ProgLength' feature

```{r var2, eval=TRUE, echo=TRUE}

# Linear model on 'ProgLength' feature
lm_1 <- lm(Drate ~ ProgLength, data = default_rates)
glance(lm_1)

```

**Conclusion:** From the above metrics, r.squared value is 0.2786509 and p-value is zero. *ProgLength* is a good predictor.

## Model evaluation on 'SchoolType' feature

```{r var3, eval=TRUE, echo=TRUE}

# Linear model on 'SchoolType' feature
lm_2 <- lm(Drate ~ SchoolType, data = default_rates)
glance(lm_2)

```

**Conclusion:** From the above metrics, r.squared value is 0.1726394 and p-value is zero. *ProgLength* is a good predictor.

## Model evaluation on 'Num' feature

```{r var4, eval=TRUE, echo=TRUE}

# Linear model on 'Num' feature
lm_3 <- lm(Drate ~ Num, data = default_rates)
tidy(lm_3)

```

**Conclusion:** *Num* feature can be considered as a predictor.

## Model evaluation on 'Denom' feature

```{r var5, eval=TRUE, echo=TRUE}

# Linear model on 'Denom' feature
lm_4 <- lm(Drate ~ Denom, data = default_rates)
tidy(lm_4)

```

**Conclusion:** *Denom* feature can be considered as a predictor.

```{r var6, eval=FALSE, echo=FALSE}

# Linear model on 'Prate' feature
lm_5 <- lm(Drate ~ Prate, data = default_rates)
tidy(lm_5)

```

## Model evaluation on 'EthnicCode' feature

```{r var7, eval=TRUE, echo=TRUE}

# Linear model on 'EthnicCode' feature
lm_6 <- lm(Drate ~ EthnicCode, data = default_rates)
glance(lm_6)

```
**Conclusion:** From the above metrics, r.squared value is 0.03850229. *EthnicCode* is a good predictor.

# Final Linear Model

## Multiple Linear Regression

```{r var8, eval=TRUE, echo=TRUE}

# Final model usiing all features
lm_0 <- lm(Drate ~ ProgLength + SchoolType + Num + Denom + EthnicCode, data = default_rates)
summary(lm_0)
plot(lm_0)

```

